---
title: "Diatom-based Calibration Model"
author: "Vélez-Agudelo Camilo"
date: "8/3/2022"
output:
  html_notebook:
    code_folding: "show"
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(vegan)
library(rioja)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(gridExtra)
library(rstatix)
```

A series of WA and WA-PLS models were developed using the screened data set (38 sampling sites and 45 diatom taxa) to identify the best calibration model that allows reliable conductivity reconstructions in northern Patagonia.

# **Preprocessing Data** #
```{r Loading.data}

speciesRC <- readRDS("speciesRC.rds") ##Colorado River Diatom Data##
speciesRN <- readRDS("speciesRN.rds") ##Negro River Diatom Data##
speciesCH <- readRDS("speciesCH.rds") ##Chubut River Diatom Data##

names <- readRDS("names.code.rds")   ##Species name code## 

varam <- readRDS("varam.rds")
chem <- readRDS("chem.rds")          ##Hidrochemical variables Log-Transformed##
```

```{r Combining.data}
colnames(speciesRC) <- names[[1]][, 2]
colnames(speciesRN) <- names[[2]][, 2]
colnames(speciesCH) <- names[[3]][, 2]

species.complete <- Merge(
  speciesRC,
  speciesRN,
  speciesCH
)

##Filter data by inland species##
species.inland <- species.complete[-c(1:3, 16:18, 34:38),]
species.inland <- species.inland[, colSums(species.inland) > 0]

##Hydochemical variables in inland sites##
chem.inland <- chem[-c(1:3, 16:18, 34:38),]
```


For comparison purposes, models with Hellinger, square-root and logarithmic transformation were applied respectively, to develop a calibration model of conductivity.

```{r Diatom.data.transformation}

species.inland.hel <- decostand(species.inland, "hellinger")
species.inland.sqrt <- sqrt(species.inland)
species.inland.log <- log1p(species.inland)
```

# **Weighted Averaging** #
```{r WA}
mod1 <- WA(
  species.inland.hel, ##Hellinger tranformed##
  chem.inland$Cond,
  tolDW=TRUE
  )

mod2 <- WA(
  species.inland.sqrt, ##Square-root tranformed##
  chem.inland$Cond,
  tolDW=TRUE
  )

mod3 <- WA(
  species.inland.log, ##Log-tranformed##
  chem.inland$Cond,
  tolDW=TRUE
  )
```

```{r WA.cross.hell1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod1.xv1 <- crossval(mod1, cv.method = "loo")
```

```{r}
mod1.xv1
```

```{r WA.cross.hell2, echo = TRUE, results = FALSE}
##Cross-validation using bootstraping##
mod1.xv2 <- crossval(mod1, cv.method = "boot", nboot = 1000)
```

```{r}
mod1.xv2
```

```{r WA.cross.square1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod2.xv1 <- crossval(mod2, cv.method = "loo")
```

```{r}
mod2.xv1
```

```{r WA.cross.square2, echo = TRUE, results = FALSE}
##Cross-validation using bootstraping##
mod2.xv2 <- crossval(mod2, cv.method = "boot", nboot = 1000) 
```

```{r}
mod2.xv2
```

```{r WA.cross.log1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod3.xv1 <- crossval(mod3, cv.method = "loo")
```

```{r}
mod3.xv1
```

```{r WA.cross.log2, echo = TRUE, results = FALSE}
##Cross-validation using Boostraping##
mod3.xv2 <- crossval(mod3, cv.method = "boot", nboot = 1000)
```

```{r}
mod3.xv2
```

# **Weighted Averaging - Partial Lineal Square** #
```{r WAPLS}
mod4 <- WAPLS(
  species.inland.hel, ##Hellinger tranformed##
  chem.inland$Cond,
  )

mod5 <- WAPLS(
  species.inland.sqrt, ##Square-root tranformed##
  chem.inland$Cond,
  )

mod6 <- WAPLS(
  species.inland.log, ##Log-tranformed##
  chem.inland$Cond,
  )
```

```{r WAPLS.hel1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod4.xv1 <- crossval(mod4, cv.method = "loo")
```

```{r}
mod4.xv1
```

```{r WAPLS.hel2, echo = TRUE, results = F}
##Cross-validation using boostraping##
mod4.xv2 <- crossval(mod4, cv.method = "boot", nboot = 1000)
```

```{r}
mod4.xv2
```

```{r WAPLS.sqrt1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod5.xv1 <- crossval(mod5, cv.method = "loo")
```

```{r}
mod5.xv1
```

```{r WAPLS.sqrt2, echo = TRUE, results = F}
##Cross-validation using boostraping##
mod5.xv2 <- crossval(mod5, cv.method = "boot", nboot = 1000)
```

```{r}
mod5.xv2
```

```{r WAPLS.log1, echo = TRUE, results = FALSE}
##Cross-validation using leave-one-out##
mod6.xv1 <- crossval(mod6, cv.method = "loo")
```

```{r}
mod6.xv1
```

```{r WAPLS.log2, echo = TRUE, results = F}
##Cross-validation using boostraping##
mod6.xv2 <- crossval(mod6, cv.method = "boot", nboot = 1000)
```

```{r}
mod6.xv2
```


# **Observed vs Inferred Plot** #


Comparison of measured and predicted conductivities revealed that the simple WA with inverse deshrinking, tolerance down-weighting and square-root transformation provided the best performance. This model showed the highest r2 jack (0.97), the lowest RMSEP (0.0521) and lowest maximum bias (0.1901). In contrast, models based on WA-PLS methods did not outperform the statistic parameters shown by the simple WA.

```{r, fig.align="center"}
best.model <- cbind(
  chem.inland$Cond,
  as.data.frame(mod2$fitted.values[, 3])
)

names(best.model) <- c("Observed", "Inferred")

ggplot(best.model,
       aes(x = Observed, y = Inferred)
       )+
  geom_point(
    size=3, 
    pch = 3
    )+
  geom_smooth(
    method = lm
    )+
  labs(
    x= "Observed Conductivity",
    y = "Inferred Conductivity"
    )+
  theme_minimal()
```

## Assumptions of Linear Regression ##
```{r Assumptions.Linear.Regression, fig.align="center"}

library(ggfortify)

reg <- lm(
  Inferred ~ Observed,
  data = best.model
)

autoplot(
  reg
  ) +
  geom_point(
    size = 2,
    alpha = 0.5,
    col = "blue"
    ) + 
  theme_bw()
```

